---
title: "Data Processing"
author: "Shannon Rush"
date: "July 10, 2014"
output: pdf_document
---

This document details the steps taken to process the Titanic Survival dataset obtained from [the Kaggle Titanic competition]("http://www.kaggle.com/c/titanic-gettingStarted") into the data needed for this Shiny application.

Load the raw data obtained from Kaggle
```{r load_data}
    data <- read.csv("train.csv")
```

Create new dataframe using just columns needed for the Shiny application, with tidier column names
```{r}
    new.data <- data.frame(survived=data$Survived,
                               class=data$Pclass,
                               gender=data$Sex,
                               age=data$Age)
```
Keep only observations that include data for each selected feature
```{r}
    titanic.data <- new.data[complete.cases(new.data),]
```
Fractional ages are not necessary for this analysis, round ages to nearest integer
```{r}
    titanic.data$age <- round(titanic.data$age)
```
Revalue survival factor for easier labeling
```{r}
    library(plyr)
    titanic.data$survived <- as.character(titanic.data$survived)
    titanic.data$survived <- revalue(titanic.data$survived, c("0"="Did Not Survive","1"="Survived"))
```

Save as .RData file for easy loading
```{r}
    save(titanic.data, file="titanic_data.RData")
```




